The Multi-Stencil Framework: A Component-Based Approach to Orchestrate Stencils

### 2017-03-06 ARTICLES/IJPP/articleR1/msf.tex

Figure~\ref{fig:msf} gives an overview of the Multi-Stencil Framework that is entirely detailed throughout this paper. It is composed of four distinct parts described hereafter.
As illustrated in Figure~\ref{fig:msf}, MSF targets two different kinds of end-users: the \emph{numerician}, in other words the mathematician, and the \emph{developer}. Most of the time numericians do have programming knowledge, however as it is not their core domain and because of a lack of time, development is often left to engineers according to numerician needs. MSF has the interesting particularity to propose a clear separation of concerns between these two end-users by distinguishing the description language from the numerical code. \CP{pas top clair la fin de la phrase}

@@

\begin{figure}[t]
\begin{center}
  \includegraphics[width=.6\textwidth]{./images/msf.pdf}
  \caption{The Multi-Stencil Framework (MSF) composed of the Multi-Stencil Language (MSL), the Generic Assembly (GA), the Multi-Stencil Compiler (MSC) and which produces a specialized assembly of components.}
  \label{fig:msf}
\end{center}
\end{figure}

\paragraph{\textbf{Multi-Stencil Language}}
The Multi-Stencil Language, or MSL, is the domain specific language proposed by the framework for the numerician. It is a descriptive language, easy to use, without any concern about implementation details.
It fits the need of a mathematician to describe the simulation. The description written with MSL can be considered as an input of the framework. MSL is described in details in Section~\ref{sect:msl}. The language is built upon the meta-model described in Section~\ref{sect:formalism}.

\paragraph{\textbf{Generic Assembly}}
In addition to the language MSL and the simulation description written by the numerician, MSF needs a Generic Assembly (GA) of a multi-stencil program as input. What is called a GA is a component assembly for which meta-types are represented and for which some parts need to be generated or specialized. A GA could be compared to a generic template or a skeleton respectively in usual object programming languages such as C++ or in functional languages. From this generic assembly will be built the final specialized assembly of the simulation where component types will be specified, and some parts of GA will be transformed or added. As well as MSL, this generic assembly is described in Section~\ref{sect:msl} and is built upon the meta-model described in Section~\ref{sect:formalism}.

@@

\paragraph{\textbf{Specialized assembly}}
Finally, the output of MSF is the component assembly generated by MSC. It is an instanciation and a transformation of the generic component assembly, by adding component types, transforming some part of the assembly, and by adding specific components generated by MSC. From this final component assembly which is specific to the simulation initially described with MSL, the developer will finally write numerical codes. This final specialized component assembly represents the parallel computations orchestration of the simulation initially described by the numerician.
\CP{Dire que c'est du \llc?}

>>>>

Figure~\ref{fig:msf} gives an overview of the Multi-Stencil Framework that is entirely detailed throughout this paper. It is composed of four distinct parts described hereafter.
As illustrated in Figure~\ref{fig:msf}, MSF targets two different kinds of end-users: the \emph{numerician}, in other words the mathematician, and the \emph{developer}. Most of the time numericians do have programming knowledge, however as it is not their core domain and because of a lack of time, development is often left to engineers according to numerician needs. MSF has the interesting particularity to propose a clear separation of concerns between these two end-users by distinguishing the description of the simulation from the implementation of numerical codes.

@@

\begin{figure}[t]
\begin{center}
  \includegraphics[width=.6\textwidth]{./images/msf.pdf}
  \caption{The Multi-Stencil Framework (MSF) is composed of the Multi-Stencil Language (MSL), the Generic Assembly (GA) and the Multi-Stencil Compiler (MSC) to produce a specialized assembly of components. The numerician, or mathematician uses MSL to describe its simulation. The developer will implement components responsible for numerical codes. A third party HPC specialist can interact with MSF to propose different version of HPC components.}
  \label{fig:msf}
\end{center}
\end{figure}

\paragraph{\textbf{Multi-Stencil Language}}
The Multi-Stencil Language, or MSL, is the domain specific language proposed by the framework for the numerician. It is a descriptive language, easy to use, without any concern about implementation details.
It fits the need of a mathematician to describe the simulation. The description written with MSL can be considered as an input of the framework. MSL is described in details in Section~\ref{sect:msl}. The language is built upon the formalism described in Section~\ref{sect:formalism}.

\paragraph{\textbf{Generic Assembly}}
In addition to the language MSL, used by the numerician to describe its simulation, MSF needs a Generic Assembly (GA) of a multi-stencil program as input. What is called a GA is a component assembly for which meta-types of components are represented and for which some parts need to be generated or specialized. A GA could be compared respectively to a template or a skeleton in object programming languages (such as C++) or functional languages. From this generic assembly will be built the final specialized assembly of the simulation where component types will be specified, and where parts of GA will be transformed. As well as MSL, this generic assembly is described in Section~\ref{sect:msl} and is built upon the meta-formalism described in Section~\ref{sect:formalism}.

@@

\paragraph{\textbf{Specialized assembly}}
Finally, the output of MSF is the component assembly generated by MSC. It is an instantiation and a transformation of the generic component assembly, by adding component types, transforming some part of the assembly, and by adding specific components generated by MSC. From this final component assembly which is specific to the simulation initially described with MSL, the developer will finally write components associated to numerical codes, or directly re-use existing components from other simulations. This final specialized component assembly is a parallel orchestration of the computations of the simulation initially described by the numerician. Finally, the specialized assembly produced by MSF is written in \llc.
